Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 2.dev

### 2024-09-17 main.tex
Bump to v2.dev.
Show todo notes, move two to issues.

\usepackage{amsmath}
\usepackage{booktabs}  % nice tables
\usepackage[margin=25mm]{geometry}
\usepackage[natbib=true,style=authoryear]{biblatex}
% NOTE : this file is automatically generated from Zotero, do not edit
% manually!
\addbibresource{references.bib}
\usepackage{siunitx}  % use for all units
\usepackage{subcaption}  % for subfigures
\usepackage[disable]{todonotes}

@@

\title{Automatic Bicycle Balance Assistance Reduces Probability of Falling at
Low Speeds When Subjected to Handlebar Perturbations\\Version 1}

@@

\subsection{Balance Assist Control}
%
We use a forward speed \(v\) gain scheduled proportional roll rate feedback
controller to stabilize the bicycle. In the speed range tested, the commanded
steer torque \(T^\textrm{m}_\delta\) from the steer motor follows the control
law
%
\begin{align}
  T^\textrm{m}_\delta = -k_{\dot{\phi}}\dot{\phi} = g(v_{stable} - v)\dot{\phi}
  \label{eq:implemented-controller}
\end{align}
%
where \(v_{\textrm{stable}} = 4.7~\si{\meter\per\second}\) is approximately the
average stable speed predicted from the open loop bicycle rigid-rider dynamics.
We use \(g=8\) and \(g=10\) as gain values during the experiments. Scaling the
proportional feedback gain linearly with respect to speed stabilizes the
normally unstable weave mode of the bicycle down to
2.9~\si{\kilo\meter\per\hour} for the riderless bicycle and
4.6~\si{\kilo\meter\per\hour} for the ridden bicycle  as shown in
Figure~\ref{fig:eigenvalues}.
\todo[inline]{Could add the experimental weave mode identification to this
figure and consider lowering the gain on the model to make the model match the
experimental values.}
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{figures/balance-assist-eig-vs-speeds.png}
  \caption{Uncontrolled (upper row) and controlled (lower row) root locus of
    eigenvalue components (real: solid, imaginary: dashed) plotted versus speed
    for the bicycle without the inertial effect of a rigid rider (left) and
    with a rigid rider (right). Vertical blue lines indicate the two speeds we
    perform experiments at. Grey shaded region is the linear stable speed
    range. Geometric and inertial parameters for these plots were estimated
    with the methods presented in \cite{Moore2012} and software packages
    BicycleParameters 1.1.1~\citep{Moore2024} and Yeadon 1.5.0~\citep{Dembia2015}.}
  \label{fig:eigenvalues}
\end{figure}

@@

\subsection{Protocol}
%
We recruited 26 able-bodied young adults (20-36 years old) to participate in
the experiments. The subjects were all confident in their cycling skills and
had cycled at least once in the last month. Eleven subjects performed the
experiments at 6~\si{\kilo\meter\per\hour} (1.7~\si{\meter\per\second}) and
fifteen subjects performed the experiments at 10~\si{\kilo\meter\per\hour}
(2.8~\si{\meter\per\second}). All subjects consented to the experiment and
could decline to continue at any time. The study was approved by Delft
University of Technology's Human Research Ethics Council (\#3897).

\todo[inline]{Include the average age, mass, height and standard deviation if
we recorded those.}

@@

\todo[inline]{Could show an impulse response in lateral deviation for different
speeds.}

>>>>

\usepackage{amsmath}
\usepackage{booktabs}  % nice tables
\usepackage[margin=25mm]{geometry}
\usepackage[natbib=true,style=authoryear]{biblatex}
% NOTE : this file is automatically generated from Zotero, do not edit
% manually!
\addbibresource{references.bib}
\usepackage{siunitx}  % use for all units
\usepackage{subcaption}  % for subfigures
\usepackage[]{todonotes}  % use [disable] to hide notes

@@

\title{Automatic Bicycle Balance Assistance Reduces Probability of Falling at
Low Speeds When Subjected to Handlebar Perturbations\\Version 2.dev}

@@

\subsection{Balance Assist Control}
%
We use a forward speed \(v\) gain scheduled proportional roll rate feedback
controller to stabilize the bicycle. In the speed range tested, the commanded
steer torque \(T^\textrm{m}_\delta\) from the steer motor follows the control
law
%
\begin{align}
  T^\textrm{m}_\delta = -k_{\dot{\phi}}\dot{\phi} = g(v_{stable} - v)\dot{\phi}
  \label{eq:implemented-controller}
\end{align}
%
where \(v_{\textrm{stable}} = 4.7~\si{\meter\per\second}\) is approximately the
average stable speed predicted from the open loop bicycle rigid-rider dynamics.
We use \(g=8\) and \(g=10\) as gain values during the experiments. Scaling the
proportional feedback gain linearly with respect to speed stabilizes the
normally unstable weave mode of the bicycle down to
2.9~\si{\kilo\meter\per\hour} for the riderless bicycle and
4.6~\si{\kilo\meter\per\hour} for the ridden bicycle  as shown in
Figure~\ref{fig:eigenvalues}.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{figures/balance-assist-eig-vs-speeds.png}
  \caption{Uncontrolled (upper row) and controlled (lower row) root locus of
    eigenvalue components (real: solid, imaginary: dashed) plotted versus speed
    for the bicycle without the inertial effect of a rigid rider (left) and
    with a rigid rider (right). Vertical blue lines indicate the two speeds we
    perform experiments at. Grey shaded region is the linear stable speed
    range. Geometric and inertial parameters for these plots were estimated
    with the methods presented in \cite{Moore2012} and software packages
    BicycleParameters 1.1.1~\citep{Moore2024} and Yeadon 1.5.0~\citep{Dembia2015}.}
  \label{fig:eigenvalues}
\end{figure}

@@

\subsection{Protocol}
%
We recruited 26 able-bodied young adults (20-36 years old) to participate in
the experiments.
\todo[]{Include the average age, mass, height and standard deviation if we
recorded those.}
The subjects were all confident in their cycling skills and had cycled at least
once in the last month. Eleven subjects performed the experiments at
6~\si{\kilo\meter\per\hour} (1.7~\si{\meter\per\second}) and fifteen subjects
performed the experiments at 10~\si{\kilo\meter\per\hour}
(2.8~\si{\meter\per\second}). All subjects consented to the experiment and
could decline to continue at any time. The study was approved by Delft
University of Technology's Human Research Ethics Council (\#3897).

@@
